Covering Your Basis: Model Agnostic Probes of the Binary Black Hole Population

### 2022-09-26 src/tex/appendicies.tex
update readme
more refs

Spline functions have been shown to be sensitivite to the chosen number of knots, and their locations or spacing. 
Adding more knots knots increases the a-priori variance in the spline function, while the space between knots can limit the 
resolution of features in the data the spline is capable of resolving. To ensure your spline based model is flexible enough 
one would want to add as many knots as densely as possible, but this comes with the inherit larger variance as mentioned above. 
To overcome this we resort to the frequentist penalized spline (P-Spline) in which one applies a penalty or regularization term 
to the likelihood based on the difference of adjacent knot coeficients. The linear combination of spline basis components is 
is flat when the basis coeficients are equal (see Figure \ref{fig:spline_basis}). By penalizing the likelihood as the differences between adjacent
knot coeficients get larger, one gets a smoothing of the spline function inferences. Since we are using hierarchical Bayesian inferece as our 
statistical framework, we formulate this penalized likelihood with its Bayesian analog. The difference penalty is equivalent to applying a stochastic 
differnece prior among the knots, as described in \bruce{CITE THIS PAPER}. This is also sometimes referred to as a Gaussian random walk prior, since
each coeficient is normally distributed about its neighboring coeficients. For a spline basis with $n$ degree's of freedom, 
and a difference penalty of order of $r$ (see \bruce{cite P-splines paper}), our prior is defined as:

@@

\noindent Above $\bm{D}_r$ is the order-$r$ difference matrix, of shape $(n-r \times n)$ and $\Gamma(a,b)$ the Gamma distribution with shape parameters, $a$ and $b$. 
This smoothing prior removes the strong dependence on number and location of knots that arises with using splines. By adding in a very large number 
of knots such that your domain is densly populated with basis coeficients, this allows the freedom for the model to find the smoothing 
scale that the data prefers. One downside of this prior specification as noted by\ bruce{CITE BAYESIAN PSPLINES}, is a slight dependence on choice 
of the b hyper-parameter for Gamma distribution prior on $\tau_\lambda$. We reduce this affect by following the method introduced in \bruce{CITE BAYESIAN PSPLINES} 
to alleviate this. This is done by modeling $\tau_\lambda$ with a mixture model on gamma distributions with varying b hyper-parameters. 
We now replace the prior on $\tau_\lambda$ with:

@@

This prior is imparting a natural attraction of the coeficients closer to each other in order to smooth the spline function, so one 
must ensure that the spline function is in fact flat given all equal coeficients. Since one needs to define n+k knots to construct an order-k basis with 
n degrees of freedom, people often will place knots on top of each other at hard parameter boundaries \bruce{cite msplines / o-splines}. 
This may seemed motivated, especially if there is no reality in the parameter beyond the boundary, but this violates the above condition. 
We follow the distinction from \bruce{CITE psplines} that the difference penalty or smooothing prior is only valid with ``proper'' spline bases. 
A proper basis is where all n+k knots are evenly spaced, see Figure \ref{fig:spline_basis}, as opposed to stacking them at the bounds.

@@

We explicitly enumerate each of the models used in this work for $p(\theta|\Lambda)$, along with 
their respective hyper-parameters and prior distributions in the next section. To calculate draw 
samples of the hyper parameters from the hierarchical posterior distribution shown in equation \ref{importance-posterior}, we use the 
NUTS hamiltonian monte carlo sampler in \textsc{numpyro} and \textsc{jax} to calculate likelihoods \bruce{FIND citations for these}.

@@

%\section{Validation Studies} \label{sec:validation}

>>>>

Spline functions have been shown to be sensitivite to the chosen number of knots, and their locations or spacing \citep{deBoor78}. 
Adding more knots knots increases the a-priori variance in the spline function, while the space between knots can limit the 
resolution of features in the data the spline is capable of resolving. To ensure your spline based model is flexible enough 
one would want to add as many knots as densely as possible, but this comes with the inherit larger variance as mentioned above. 
To overcome this we resort to the frequentist penalized spline (P-Spline) in which one applies a penalty or regularization term 
to the likelihood based on the difference of adjacent knot coeficients \citep{eilers2021practical}. The linear combination of spline basis components 
or the resulting spline function is flat when the basis coeficients are equal (see Figure \ref{fig:spline_basis}). By penalizing the likelihood as 
the differences between adjacent knot coeficients get larger, one gets a smoothing of the spline function inferences \citep{eilers2021practical}. 
Since we are using hierarchical Bayesian inferece as our statistical framework, we formulate this penalized likelihood with its Bayesian analog. 
The difference penalty is equivalent to applying a stochastic differnece prior among the knots, as described in  \\citet{BayesianPSplines,Jullion2007RobustSO}. 
This is also sometimes referred to as a Gaussian random walk prior, since each coeficient is normally distributed about its neighboring coeficients. 
For a spline basis with $n$ degree's of freedom, and a difference penalty of order of $r$ (see \citet{eilers2021practical}), the smoothing prior is defined as:

@@

\noindent Above $\bm{D}_r$ is the order-$r$ difference matrix, of shape $(n-r \times n)$ and $\Gamma(a,b)$ the Gamma distribution with shape parameters, $a$ and $b$. 
This smoothing prior removes the strong dependence on number and location of knots that arises with using splines. By adding in a very large number 
of knots such that your domain is densly populated with basis coeficients, this allows the freedom for the model to find the smoothing 
scale that the data prefers. One downside of this prior specification as noted in \citet{BayesianPSplines} and \citet{Jullion2007RobustSO}, is a slight dependence on choice 
of the b hyper-parameter for Gamma distribution prior on $\tau_\lambda$. We reduce this affect by following the method introduced in \citet{Jullion2007RobustSO}
to alleviate this. This is done by modeling $\tau_\lambda$ with a mixture model on gamma distributions with varying b hyper-parameters. 
We now replace the prior on $\tau_\lambda$ with:

@@

This prior is imparting a natural attraction of the coeficients closer to each other in order to smooth the spline function, so one 
must ensure that the spline function is in fact flat given all equal coeficients. Since one needs to define n+k knots to construct an order-k basis with 
n degrees of freedom, people often will place knots on top of each other at hard parameter boundaries \cite{deBoor78,monotone_regression_splines}. 
This may seemed motivated, especially if there is no reality in the parameter beyond the boundary, but this violates the above condition. 
We follow the distinction from \citet{eilers2021practical} that the difference penalty or smooothing prior is only valid with ``proper'' spline bases. 
A proper basis is where all n+k knots are evenly spaced, see Figure \ref{fig:spline_basis}, as opposed to stacking them at the bounds.

@@

We explicitly enumerate each of the models used in this work for $p(\theta|\Lambda)$, along with 
their respective hyper-parameters and prior distributions in the next section. To calculate draw 
samples of the hyper parameters from the hierarchical posterior distribution shown in equation \ref{importance-posterior}, we use the 
NUTS hamiltonian monte carlo sampler in \textsc{NumPyro} and \textsc{Jax} to calculate likelihoods \citep{jax,pyro,numpyro}.

@@

\section{Validation Studies} \label{sec:validation}

\bruce{Need to fill this section}
